Local to Global Mariculture Modelling

### 2025-07-23 06_run_atlantic_salmon_model.qmd
Farm outputs processing

targets::tar_validate()
# targets::tar_visnetwork(targets_only = T)
# targets::tar_outdated()
targets::tar_make(
  names = "sens_results_spec",
  seconds_meta_append = 120
  )
```

@@

Note that this also takes a long time - each branch of farmrun takes ~4.5 minutes, and a maximum of ~12 workers can run at a time. The total therefore takes ~17 hours.

@@

targets::tar_validate()
# targets::tar_visnetwork(targets_only = T)
# targets::tar_outdated()
targets::tar_make(farm_results, seconds_meta_append = 90)
meta <- targets::tar_meta() %>% 
  filter(grepl("farm_run", name))
meta$seconds %>% mean()/60
```

```{r}
#| label: farm-run-results

targets::tar_load(cohort_results)
```

>>>>

targets::tar_validate()
# targets::tar_visnetwork(targets_only = T)
# targets::tar_outdated()
targets::tar_make(
  seconds_meta_append = 120
  )
```

@@

Note that this takes a long time - each branch of farmrun takes ~5 minutes, and a maximum of ~12 workers can run at a time. The total therefore takes ~17 hours.

@@

targets::tar_validate()
# targets::tar_visnetwork(targets_only = T)
# targets::tar_outdated()
targets::tar_prune()

targets::tar_make(seconds_meta_append = 90)

meta <- targets::tar_meta() %>% 
  filter(grepl("farm_run", name))
meta$seconds %>% mean()/60 # minutes per target
```

```{r}
#| label: all-farms-cohorts-results

targets::tar_load(stat_names)
targets::tar_load(farm_results)
targets::tar_load(cohort_results)

purrr::map(stat_names, function(stat_name) {
  farm_results %>% 
    filter(measure == stat_name) %>% 
    qsave(file.path(output_model_farm_path, str_c(stat_name, "_all_farms.qs")))

  cohort_results %>% 
    filter(measure == stat_name) %>% 
    qsave(file.path(output_model_cohort_path, str_c(stat_name, "_all_farms.qs")))
})

rm(farm_results, cohort_results)
```

```{r}
#| label: per-biomass-ouputs

farm_fnms <- list.files(output_model_farm_path, full.names = T)
cohort_fnms <- list.files(output_model_cohort_path, full.names = T)

farm_bioms <- farm_fnms %>% 
  str_subset("biomass") %>% 
  qs::qread() %>% 
  distinct(farm_ID, feed, t, mean, sd)
cohort_bioms <- cohort_fnms %>% 
  str_subset("biomass") %>% 
  qs::qread() %>% 
  distinct(farm_ID, feed, t, mean, sd)

stat_names <- stat_names[!stat_names %in% c("biomass_stat", "weight_stat", "dw_stat", "water_temp_stat", "T_response_stat", "rel_feeding_stat")]

purrr::map(stat_names, function(stat_name) {
  farm_fnms %>% 
    str_subset(stat_name) %>% 
    qs::qread() %>% 
    merge(farm_bioms, by = c("farm_ID", "feed", "t")) %>% 
    rename(
      mean = mean.x,
      sd = sd.x,
      biom_mean = mean.y,
      biom_sd = sd.y
    ) %>% 
    mutate(
      biom_sd_mean = biom_sd/biom_mean,
      mean_perbiom = mean/biom_mean,
      sd_perbiom = (sd_mean+biom_sd_mean) * mean_perbiom
    ) %>% 
      qs::qsave(file.path(per_biomass_farm_path, str_c(stat_name, "_all_farms.qs")))

  cohort_fnms %>% 
    str_subset(stat_name) %>% 
    qs::qread() %>% 
    merge(cohort_bioms, by = c("farm_ID", "feed", "t")) %>% 
    rename(
      mean = mean.x,
      sd = sd.x,
      biom_mean = mean.y,
      biom_sd = sd.y
    ) %>% 
    mutate(
      biom_sd_mean = biom_sd/biom_mean,
      mean_perbiom = mean/biom_mean,
      sd_perbiom = (sd_mean+biom_sd_mean) * mean_perbiom
    ) %>% 
      qs::qsave(file.path(per_biomass_cohort_path, str_c(stat_name, "_all_farms.qs")))
})
```

### 2025-07-23 index.qmd
Farm outputs processing

```{r global-variables}
#| include: false

@@

feed_pal <- c("past" = "#E41A1C", "reference" = "#377EB8", "future" = "#4DAF4A")  
cohort_pal <- c("1" = "#E41A1C", "2" = "#377EB8", "3" = "#4DAF4A", "all" = "black")
macro_pal <- c(
  "protein" = "darkred", "P" = "darkred", 
  "carbohydrates" = "steelblue", "C" = "steelblue", 
  "lipids" = "darkgoldenrod1", "L" = "darkgoldenrod1"

@@

```{r raw biomass and weight data}
weight_cohort <- find_read(output_model_cohort_path, "weight_stat_all_farms")
biomass_cohort <- find_read(output_model_cohort_path, "biomass_stat_all_farms")
```
)
```

>>>>

```{r global-variables}
#| include: false
#| cache: false

@@

feed_pal <- c(
  "past" = "#E41A1C", 
  "reference" = "#377EB8", 
  "future" = "#4DAF4A",
  "plant_dominant" = "#4DAF4A",
  "marine_dominant" = "#377EB8",
  "novel_inclusive" = "#E41A1C"
)
cohort_pal <- c("1" = "#E41A1C", "2" = "#377EB8", "3" = "#4DAF4A", "all" = "black")
macro_pal <- c(
  "protein" = "darkred", 
  "P" = "darkred", 
  "carbohydrates" = "steelblue", 
  "C" = "steelblue", 
  "lipids" = "darkgoldenrod1", 
  "L" = "darkgoldenrod1"
)

prettyplot <- function() {
  theme_classic() +
    theme(
      text = element_text(family = "sans", size = 12, colour = "black"),
      legend.position = "none"
    )
}
```

@@

```{r raw biomass and weight data}
weight_cohort <- find_read(output_model_cohort_path, "weight_stat_all_farms")
biomass_cohort <- find_read(output_model_cohort_path, "biomass_stat_all_farms")
```

Each farm's final population was calculated by dividing the production tonnage required by the average modelled individual harvest size for each farm (when feed reference feed, @fig-harvest-size). A population timeseries was then generated using the standard mortality rate (`r round(unname(pop_params['mortmyt'])*100, 3)` % d$^{-1}$).

```{r}
#| label: fig-harvest-size
#| fig-cap: |
#|   Frequency distribution of harvest sizes used to create the population timeseries for each farm, as predicted by the fish-growth model.

harv_size <- find_read(output_farm_data_path, "farm_harvest_size")

harv_size %>% 
  mutate(weight = weight %>% set_units("g") %>% set_units("kg")) %>% 
  select(c(farm_ID, weight)) %>% 
  ggplot(aes(x = weight)) +
  geom_histogram(binwidth = 0.25, colour = "black", fill = "salmon", alpha = 0.75) +
  scale_y_continuous(limits = c(0,500)) +
  labs(y = "Frequency", x = "Harvest weight") +
  prettyplot() +
  theme(aspect.ratio = 0.75)
```

Using this method, a farm's 'population' is not a good representation of the actual number of individuals being farmed, but it ensures each farm location is producing the correct tonnage while allowing comparisons between the different feeds.

```{r}
#| label: fig-cohorts
#| eval: false
#| fig-cap: | 
#|   Example of how the overlapping cohorts combine into a single farm production curve (in this case biomass). The left plot shows the farm biomass produced by each cohort from initialisation to harvest, while the right plot shows the combined farm biomass over the course of a single production period of 548 days.

fnms <- output_growth_data_path %>% 
  list.files(full.names = T) %>% 
  str_subset("reference")

df <- qs::qread(fnms[1])[["biomass_stat"]] %>% 
  as.data.frame() %>% 
  mutate(mean = set_units(mean, "g") %>% set_units("t"),
         sd = set_units(sd, "g") %>% set_units("t")) %>% 
  select(-farm_ID)

df <- rbind(
  df %>% mutate(cohort = 1),
  df %>% mutate(t = t+365, cohort = 2),
  df %>% mutate(t = t+730, cohort = 3)
) %>% 
  mutate(t = (t-min(t))+1,
         split = "Split by cohort")

min_t <- min(df$t[df$cohort == 2])
max_t <- max(df$t[df$cohort == 2])

df_2 <- df %>% 
  filter(t %in% min_t:max_t) %>% 
  mutate(sd = sd/mean) %>% 
  group_by(t) %>% 
  reframe(sd = sqrt(sum(sd^2)),
          mean = sum(mean),
          sd = sd*mean) %>% 
  mutate(cohort = "all",
         split = "Totalled")

df <- rbind(df, df_2) %>% 
  mutate(cohort = as.factor(cohort))

ggplot(df, aes(x = t, y = mean, ymin = mean-sd, ymax = mean+sd, colour = as.factor(cohort))) +
  geom_line(linewidth = 0.75) +
  geom_ribbon(linetype = "dotted", alpha = 0) +
  facet_grid(cols = vars(split)) +
  scale_x_continuous(breaks = seq(0, 1600, 120)) +
  scale_colour_manual(values = cohort_pal) +
  prettyplot() +
  labs(y = "Farm biomass", x = "Day of production")
```

```{r}
#| label: pop-biom-data

biomass_farm <- find_read(output_model_farm_path, "biomass_stat") %>% 
  group_by(farm_ID, feed) %>% 
  slice_max(t) %>% 
  ungroup()

production <- find_read(output_farm_data_path, "farm_static_data") %>% 
  merge(biomass_farm, by = "farm_ID", all.x = F) %>% 
  mutate(
    mean = set_units(mean, "g"),
    diff = set_units(mean - tonnes_per_farm, "t"),
    perc_diff = diff/tonnes_per_farm
  ) %>% 
  group_by(country, feed) %>% 
  reframe(
    min_percdiff = min(perc_diff),
    max_percdiff = max(perc_diff),
    mean_percdiff = mean(perc_diff)
  )
```

By this method, farm biomass as predicted by the model predicted farm tonnage relatively accurately. Predicted farm biomass was within `r round(100*min(production$min_percdiff),1)`% -- `r round(100*max(production$max_percdiff),1)`% of tonnage produced.

## Feed provision and feeding response to temperature {#sec-model-feeding}

# Modelled experimental feeds {#sec-feeds}

# Model sensitivity to parameter changes {#param-sensitivity}

Parameter sensitivity was tested by varying each parameter by $\pm 10$% and calculating sensitivity as $S_p = \frac{(P_{1.1} - P_{0.9})}{0.2 \times P_1}$, where $S$ is the sensitivity of the measured variable to changes in parameter $p$, and $P_{1.1}$, $P_{0.9}$, and $P_{1}$ are the value of the measured variable when the model is run with parameter $p$ at 110%, 90%, and 100% of its value respectively. 
All parameters in Table @tbl-params were tested. The population-level parameters ($\overline{W}$, $\Delta{W}$, $\overline{I_{max}}$, $\Delta{I_{max}}$, $\overline{F}$, $\Delta{F}$) were tested with a population of of 5000 on a random 10% of farms, while all other parameters were tested across all farms with a population of 10.

```{r salmon-param-names}
#| code-summary: Get parameter and sensitivity data

params <- file.path(output_species_data_path, "sens_params.qs") %>% qs::qread()

param_names <- tibble::tribble(
  ~name,        ~lab,
  "alpha",      bquote(alpha),
  "epsprot",    bquote(epsilon["P"]),
  "epslip",     bquote(epsilon["L"]),
  "epscarb",    bquote(epsilon["C"]),
  "epsO2",      bquote(epsilon["O"[2]]),
  "pk",         bquote("pk"),
  "k0",         bquote("k"[0]),
  "m",          bquote("m"),
  "n",          bquote("n"),
  "betac",      bquote(beta*"C"),
  "Tma",        bquote("T"["max"]),
  "Toa",        bquote("T"["opt"]),
  "Taa",        bquote("T"["min"]),
  "omega",      bquote(omega),
  "a",          bquote("a"),
  "k",          bquote("k"),
  "eff",        bquote("eff"),
  "meanW",      bquote(bar("W")),
  "deltaW",     bquote(Delta*"W"), 
  "meanImax",   bquote(bar("I"["max"])),
  "deltaImax",  bquote(Delta*"I"["max"]),
  "overFmean",  bquote(bar("F")),
  "overFdelta", bquote(Delta*"F"),
  "mortmyt",    bquote("mort")
)
```

```{r}
#| label: sens-results

sens_results_files <- file.path(output_sens_data_path) %>% 
  list.files(full.names = T) %>% 
  str_subset("results")
```

# Inputs {#inputs}

```{r}
#| label: farm-geometry

farm_geometry <- find_read(output_farm_data_path, "farm_geometry")
```

```{r}
#| label: input-data

ins_P <- list.files(per_biomass_cohort_path, full.names = T) %>% 
  str_subset("P_uneat|P_excr") %>% 
  purrr::map(function(nm) {qs::qread(nm)}) %>% 
  bind_rows()

ins_tot <- list.files(per_biomass_cohort_path, full.names = T) %>% 
  str_subset("total_uneat|total_excr") %>% 
  purrr::map(function(nm) {qs::qread(nm)}) %>% 
  bind_rows()
```

## Total inputs (uneaten and excreted)

### Through time

```{r}
#| label: fig-uneat-excr-plot-ts
#| fig-cap: Initial look at excreted and uneaten through time

ins_tot %>% 
  mutate(t = t-365) %>% 
  merge(farm_geometry, by.x = "farm_ID", by.y = "farm_id") %>% 
  group_by(country, feed, t, measure) %>% 
  reframe(
    sd_perbiom = sd(mean_perbiom),
    mean_perbiom = mean(mean_perbiom)
    ) %>% 
  mutate(
    mean_perbiom = mean_perbiom %>% 
      set_units("g g_fish-1") %>% 
      set_units("kg t_fish-1"),
    sd_perbiom = sd_perbiom %>% 
      set_units("g g_fish-1") %>% 
      set_units("kg t_fish-1")
  ) %>% 
  ggplot(aes(
    x = t, 
    y = drop_units(mean_perbiom), 
    ymin = drop_units(mean_perbiom-sd_perbiom), 
    ymax = drop_units(mean_perbiom+sd_perbiom), 
    colour = feed, 
    fill = feed
  )) +
  geom_line() +
  geom_ribbon(alpha = 0.25) +
  facet_grid(
    rows = vars(country),
    cols = vars(measure),
    scales = "free"
  ) +
  prettyplot() +
  scale_colour_manual(values = feed_pal) +
  scale_fill_manual(values = feed_pal) +
  scale_y_continuous(
    name = expression("Loss (kg g"^-1*")"),
    breaks = seq(0, 1, 0.2)
  )
```

```{r}
#| label: uneat-excr-plot-perc-diff

tmp <- ins_tot %>% 
  mutate(t = t-365) %>% 
  merge(farm_geometry, by.x = "farm_ID", by.y = "farm_id") %>% 
  dplyr::select(-c(mean, sd_mean, sd, biom_mean, biom_sd, biom_sd_mean, sd_perbiom)) %>% 
  group_by(country, t, measure) %>% 
  pivot_wider(
    names_from = feed, 
    values_from = mean_perbiom
  ) %>% 
  mutate(
    diff = plant_dominant - marine_dominant,
    perc_diff = diff/plant_dominant
  ) %>%
  group_by(country, t, measure) %>% 
  reframe(
    mean_diff = mean(diff),
    mean_perc_diff = mean(perc_diff),
    sd_diff = sd(diff),
    sd_perc_diff = sd(perc_diff)
  )

tmp <- tmp %>% 
  group_by(measure, country) %>% 
  reframe(
    mean_t_perc_diff = mean(mean_perc_diff),
    sd_t_perc_diff = sd(mean_perc_diff)
    )
```

While the actual difference in excreted and uneaten feed (total loss) varied between countries and through time (@fig-uneat-excr-plot-ts), relative changes between the feeds remained consistent. 
In both the US and Australia (all done so far), total excreted faeces was 
`r round(100*tmp$mean_t_perc_diff[tmp$country == "Australia" & tmp$measure == "total_excr_stat"], 1)`% ($\pm$ `r round(100*tmp$sd_t_perc_diff[tmp$country == "Australia" & tmp$measure == "total_excr_stat"], 1)`%) higher in the plant-dominant feed. 
Differences in total uneaten feed were much smaller at 
`r round(100*-tmp$mean_t_perc_diff[tmp$country == "Australia" & tmp$measure == "total_uneat_stat"], 1)`% ($\pm$ `r round(100*tmp$sd_t_perc_diff[tmp$country == "Australia" & tmp$measure == "total_uneat_stat"], 1)`%) lower in the plant-dominant feed.

### At harvest

```{r}
#| label: uneat-excr-plot-end
#| fig-cap: Caption

ins_tot %>% 
  group_by(farm_ID) %>%
  arrange(farm_ID, t) %>%
  mutate(
    cumsum_mean = cumsum(mean_perbiom) %>% 
      set_units("g g_fish-1") %>% 
      set_units("kg t_fish-1"),
    end_t = max(t)
    ) %>% 
  ungroup() %>% 
  filter(t == end_t) %>% 
  merge(farm_geometry, by.x = "farm_ID", by.y = "farm_id") %>% 
  ggplot(aes(x = country, y = cumsum_mean, fill = feed)) +
  geom_boxplot() +
  facet_grid(cols = vars(measure)) +
  prettyplot() +
  scale_fill_manual(values = feed_pal)
```

## N inputs

```{r}
#| label: N-loss-plot-total
#| fig-cap: Caption

ins_P %>% 
  group_by(farm_ID) %>%
  arrange(farm_ID, t) %>%
  mutate(
    cumsum_mean = cumsum(mean_perbiom)*6.25 %>% 
      set_units("g g_fish-1") %>% 
      set_units("kg t_fish-1"),
    end_t = max(t)
    ) %>% 
  ungroup() %>% 
  filter(t == end_t) %>% 
  merge(farm_geometry, by.x = "farm_ID", by.y = "farm_id") %>% 
  ggplot(aes(x = country, y = cumsum_mean, fill = feed)) +
  geom_boxplot() +
  facet_grid(cols = vars(measure)) +
  prettyplot() +
  scale_fill_manual(values = feed_pal)
```

# Code and packages used

All analysis was conducted in R version `r R.version[['major']]`.`r R.version[['minor']]` "`r R.version[['nickname']]`" [@R_base]. 
Major packages used include `r cite_packages(packages)`.

For a full list of R packages used see the lockfile on [Github](https://www.github.com).
This manuscript was written in Quarto [@quarto] using TinyTex [@tinytex] and the acronyms extension [@acronyms].
